### 2014-10-14 ApTranscriptome_TR.Rmd
added plot with critical temperature of up and downregulation

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods) # required by `knitcitations` when file is `knit` from command-line
library(reshape2)
library(RColorBrewer)
library(coin)

@@

Thermally-responsive genes could also differ in the temperatures at which they have increased or decreased expression in response to temperature changes. To examine this, I determine the temperature at which each responsive gene has the greated increase or decrease in expression.

@@

# make data.frame for results
l5 <- length(unique(A22.int.TPM.dt.sub$Transcript))
A22.int.T_off <- data.frame(Transcript = unique(A22.int.TPM.dt.sub$Transcript), colony = rep("ApVT", length = l5), type = rep("Low", length = l5), T_off_high = NA, T_off_low = NA)

@@

    subdf.high <- subdf[which(subdf$val >= 21), ]
    T_off_high <- subdf.high[median(which(diff(subdf.high$TPM) == min(diff(subdf.high$TPM)))), val]
    A22.int.T_off[which(A22.int.T_off$Transcript == i), "T_off_high"] <- T_off_high

    subdf.low <- subdf[which(subdf$val <= 21), ]
    T_off_low <- subdf.low[median(which(diff(subdf.low$TPM) == min(diff(subdf.low$TPM)))), val]
    A22.int.T_off[which(A22.int.T_off$Transcript == i), "T_off_low"] <- T_off_low
}

@@

# make data.frame for results
l6 <- length(unique(Ar.int.TPM.dt.sub$Transcript))
Ar.int.T_off <- data.frame(Transcript = unique(Ar.int.TPM.dt.sub$Transcript), colony = rep("ApVT", length = l6), type = rep("Low", length = l6), T_off_high = NA, T_off_low = NA)

@@

    subdf.high <- subdf[which(subdf$val >= 21), ]
    T_off_high <- subdf.high[median(which(diff(subdf.high$TPM) == min(diff(subdf.high$TPM)))), val]
    Ar.int.T_off[which(Ar.int.T_off$Transcript == i), "T_off_high"] <- T_off_high

    subdf.low <- subdf[which(subdf$val <= 21), ]
    T_off_low <- subdf.low[median(which(diff(subdf.low$TPM) == min(diff(subdf.low$TPM)))), val]
    Ar.int.T_off[which(Ar.int.T_off$Transcript == i), "T_off_low"] <- T_off_low
}

# compare T_off among species
(T_off.int.high.ttest <- t.test(Ar.int.T_off$T_off_high, A22.int.T_off$T_off_high))

(T_off.int.low.ttest <- t.test(Ar.int.T_off$T_off_low, A22.int.T_off$T_off_low))
```

As with genes turned on at high temperatures, *A. carolinensis* **Intermediate** genes are down-regulated on average at 1.2C higher temperatures than *A. picea* **Intermediate** genes.

>>>>

# Load libraries
library(R.utils)
library(ggplot2)
library(knitr)
library(knitcitations)
library(pander)
library(stringr)
library(data.table)
library(RCurl)
library(plyr)
library(MASS)
library(xtable)
library(methods) # required by `knitcitations` when file is `knit` from command-line
library(reshape2)
library(RColorBrewer)
library(coin)
library(gridExtra)

@@

Thermally-responsive genes could also differ in the critical temperature of gene induction. To examine this, I determine the temperature at which each responsive gene has the greatest increase or decrease in expression.

@@

# make data.frame for results
l5 <- length(unique(A22.int.TPM.dt.sub$Transcript))
A22.int.T_off <- data.frame(Transcript = unique(A22.int.TPM.dt.sub$Transcript), colony = rep("ApVT", length = l5), type = rep("Intermediate", length = l5), High = NA, Low = NA)

@@

    subdf.high <- subdf[which(subdf$val >= 21), ]
    High <- subdf.high[median(which(diff(subdf.high$TPM) == min(diff(subdf.high$TPM)))), val]
    A22.int.T_off[which(A22.int.T_off$Transcript == i), "High"] <- High

    subdf.low <- subdf[which(subdf$val <= 21), ]
    Low <- subdf.low[median(which(diff(subdf.low$TPM) == min(diff(subdf.low$TPM)))), val]
    A22.int.T_off[which(A22.int.T_off$Transcript == i), "Low"] <- Low
}

@@

# make data.frame for results
l6 <- length(unique(Ar.int.TPM.dt.sub$Transcript))
Ar.int.T_off <- data.frame(Transcript = unique(Ar.int.TPM.dt.sub$Transcript), colony = rep("AcNC", length = l6), type = rep("Intermediate", length = l6), High = NA, Low = NA)

@@

    subdf.high <- subdf[which(subdf$val >= 21), ]
    High <- subdf.high[median(which(diff(subdf.high$TPM) == min(diff(subdf.high$TPM)))), val]
    Ar.int.T_off[which(Ar.int.T_off$Transcript == i), "High"] <- High

    subdf.low <- subdf[which(subdf$val <= 21), ]
    Low <- subdf.low[median(which(diff(subdf.low$TPM) == min(diff(subdf.low$TPM)))), val]
    Ar.int.T_off[which(Ar.int.T_off$Transcript == i), "Low"] <- Low
}

# compare T_off among species
(T_off.int.high.ttest <- t.test(Ar.int.T_off$High, A22.int.T_off$High))

(T_off.int.low.ttest <- t.test(Ar.int.T_off$Low, A22.int.T_off$Low))
```

As with genes turned on at high temperatures, *A. carolinensis* **Intermediate** genes are down-regulated on average at 2.7°C higher temperatures than *A. picea* **Intermediate** genes. In contrast, towards low temperatures, the temperature of mean temperature of downregulation was higher for *A. picea* (10.6°) than *A. carolinensis* (9.75°).

Visualize T~off~ for both *Low* and *High* genes on the same plot.

```{r activation_plot, echo=FALSE, eval=TRUE}
# reshape data
Ar.T_off.df <- melt(Ar.int.T_off, id.vars = c("Transcript", "colony", "type"), 
                 measure.vars = c("High", "Low"),
                 variable.name = "class",
                 value.name = "T_off")

A22.T_off.df <- melt(A22.int.T_off, id.vars = c("Transcript", "colony", "type"), 
                 measure.vars = c("High", "Low"),
                 variable.name = "class",
                 value.name = "T_off")

# merge data for plotting
Ap.T_off <- rbind(Ar.T_off.df, A22.T_off.df)
Ap.T_off$Species <- ifelse(Ap.T_off$colony == "ApVT", "A. picea", "A. carolinensis")
# change factor levels to "Low" first
Ap.T_off$class = factor(Ap.T_off$class, levels = c("Low", "High"))

# combined plot
T_off_plot <- ggplot(Ap.T_off, aes(x = T_off, y = ..density.., colour = Species, group = Species, fill = Species)) +
    facet_grid(. ~ class) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    labs(x = "Temperature of gene activation", y = "Density") +
    theme(axis.title = element_text(size = rel(2))) +
    theme(axis.text = element_text(size = rel(1.2)))
#    theme(legend.position = "none")
print(T_off_plot)
```

Make a plot with all 4 gene activation comparisons.

```{r combined_plot}
gg2 <- ggplot(Ap.T_on, aes(x = T_on, y = ..density.., colour = Species, group = Species, fill = Species)) +
    facet_grid(. ~ type) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    labs(x = "Critical temperature of upregulation", y = "Density") +
    theme(legend.position = "none")

gg3 <- ggplot(Ap.T_off, aes(x = T_off, y = ..density.., colour = Species, group = Species, fill = Species)) +
    facet_grid(. ~ class) +
    geom_histogram(position = "dodge", binwidth = 3) +
    geom_density(adjust = 3, fill = NA, size = 2) + 
    labs(x = "Critical temperature of downregulation", y = "Density") +
    theme(legend.position = "bottom", legend.text = element_text(size = 12, face = 3))

# grid.arrange
grid.arrange(gg2, gg3, nrow = 2)
```
